\textbf{\sc Deliberative Autonomy for Marine Robotics}

### 2012-01-26 trunk/results.tex
updates

>>>>

\paragraph {Mixed-Initiative Autonomy}

\begin{figure*}
\centering
\subfloat[The ODSS Concept]{\label{fig:odss-concept}\includegraphics[width=0.5\textwidth]{figs/odss-concept.pdf}}\\
\subfloat[Current Implementation for \texttt{CANON}]{\label{fig:odss-frontend}\includegraphics[width=0.5\textwidth]{figs/dss-screenshot.png}}
\caption{\small The Oceanographic Decision Support System (ODSS) is a
  single portal for planning, visualization, situational awareness,
  collaboration and archival of information before, during and after
  field experiments. The ODSS has been used since mid-year 2010 for
  \texttt{CANON} field experiments.}
\label{fig:odss}
\end{figure*}

Large scale oceanographic field experiments akin to those targeted by
\can are usually logistically constrained driven largely by ship
schedules.  Once a feature set has been identified for scientific
observation and science goals have been collated, the appropriate
robotic assets need to be assigned and deployed for observation and
sampling. This asset allocation in turn depends on parameters like
sensor payload, the speed of the robot \situ in comparison to the
spatio-temporal scales of the evolution of the phenomena being
sampled, the number of assets in a class, the kinds of measurements
needed (e.g synoptic or close-up), logistical considerations for
deployments such as battery charge needed or distance/time to deploy,
impact to long-range deployment plans among others. All drive the
strategy to balance short and long-term scientific and engineering
needs. Our \can deliberations during the 2010 field season for example,
were driven by all of the above \cite{das11}. Often these strategies
were rebalanced during each decision-making cycle, especially in light
of opportunistic science goals. For example, gliders running at $0.5$
knots were often retargeted when surface support vessels noted the
apparent dispersal of the bloom to subsurface layers when HABs were a
\texttt{CANON} target. However, their retargeting was carefully
monitored to balance the long-term desire of obtaining time series
within fixed tracks and with the need to ensure these assets were not
exposed to strong currents. Powered AUVs were targeted to chase after
advecting blooms, but the decision on their deployment locations and
survey areas too close to the surf-zone were avoided. In
patch-tracking experiments, the deployment of drifters close to shore
and where depths of $20m$ or less were encountered, was avoided even
when contrary to scientific intent. The need to obtain high-quality
data for post-hoc analysis was also driven by the response time of
data returned by experiment partners.

\begin{figure*}
\centering
\subfloat[]{\label{fig:odss-mtg}\includegraphics[width=0.3\textwidth]{figs/october-canon-meeting.png}}
\subfloat[]{\label{fig:odss-virtual}\includegraphics[width=0.33\textwidth]{figs/virtual-drifter-tool.png}}
\subfloat[]{\label{fig:odss-ipad}\includegraphics[width=0.3\textwidth]{figs/photo-ipad-2.jpg}}
\caption{\small Fig. \ref{fig:odss-mtg} shows a 2010 \can meeting with the
  ODSS in use while Fig. \ref{fig:odss-ipad} shows its use on a portable
  device.}
xf\label{fig:odss2}
\end{figure*}

These distinctions make some vehicles more or less suitable to address
certain science goals and suggest some degree of intelligent planning
in the assignment strategy. Further, even if partial plans can be
synthesized automatically, the need to capture qualitative 'intent'
and desires in the final outcome of an asset allocation strategy
drives us towards a human-in-the-loop approach for asset
allocation. The vast uncertainty and complex interaction between
unknown quantities in the ocean environment dictate that any
comprehensive experimental methodology must rely in some measure on
human reasoning and adaptability.  Furthermore, the multidisciplinary
nature of ocean experiments suggests collaboration between multiple
human actors. Human cognitive capability then to capture these
uncertainties, evaluate options, factor intent and synthesize data
needed for asset allocation becomes increasingly important. Therefore,
mixed-initiative approaches are viable as a means to apportion problem
solving and inference between human and computational approaches. A
viable example was shown in the command and control of the twin Mars
rovers in \cite{aichang04,bresina05,bresina05a}\footnote{The MAPGEN
  system uses the same \eu planner used in \rx. It continues to be
  used routinely to this day on the MER mission.}.

Additionally, our work on ODSS is driven by a vision of how marine
field robotics will likely be conducted in the future and the
necessary computational infrastructure that will augment decision
making, onshore as well as on-board. We believe ultimately ODSS will
combine capabilities inferential and data gathering and analysis
capabilities and more uniquely, provide an event response capability
that could trigger the deployment of a robot and/or adapt its sampling
strategy, all from a scientist's desktop on shore. Data returned from
robotic assets at sea will be filtered for signals associated with
feature(s) of interest along with information on its contextual
environment. The association between the signal and a prospective
event will be undertaken through tagging by experts and machine
learning techniques used in recommender
systems\cite{Adomavicius05}. Adaptive sampling strategies will then
balance decision-making \situ with that done onshore with
computational models and human agents, the subject of a new NSF
project in which the authors are participating.

Operationally, the ODSS has become a software system that the
scientist consults during a \can field campaign. S/he consults
up-to-date model outputs, synoptic views of ocean color from remote
sensing data and weather conditions for the next $24$-hour planning
cycle and then queries the system to compute candidate sampling
locations. In the near future, the ODSS will consult its situational
awareness database, incorporate model outputs projecting these outputs
into the space-time region of interest to generates candidate
hypotheses with associated utility metrics. With the scientist's
guidance the automated planner would then determine the most viable
robotic option(s) to be dispatched (underwater, surface or aerial),
given available resources, distance of travel, required payload and
the environment. For instance, unpowered autonomous gliders would be
chosen for longer-duration missions with smaller payload requirements
in areas of moderate to weak ocean currents. The automated planner
would balance these criteria to compute a viable trajectory plan for
the selected vehicle. Vehicles with more controllability would be
given a specified area to survey and the path generated \situ. Goals
are communicated to the robot after analysis on shore for selective
targeting of features by the conglomerate of vehicles in the ocean
being tracked by the ODSS. Thus far, the ODSS has provided a simple
yet functional tool for scientists to use in the daily complex
decision making and the following experiment and survey design
needs. It has helped train scientists to think of expanding their
repertoire of tools for the scientific process and in the process it
enabled a mixed team of engineers and scientists to interact at the
functional level of the systems' capabilities.
